---
title: "Bachelor's Thesis: Simulation Data EDA"
author: "Andrey Peshev"
output: html_notebook
---
* prb = Percentage Relative Bias  
* mcsd = Monte Carlo Standard Deviation\
* Data generation equation: Y = 1.0 + 0.33X + 0.33Z1 + 0.33Z2 + ε\

Simulation parameters: \
Proportion of the MAR predictor represented in the analysis model: pMAR = {1.0, 0.75, 0.5, 0.25, 0.0}\
Strength of correlations among the predictors: rXZ = {0.0, 0.1, 0.3, 0.5}\
Sample size: N = {100, 250, 500}\
Proportion of missing data: PM = {0.1, 0.2, 0.4}\
R^2^ for the data generating model: R^2 = {0.15, 0.3, 0.6}\


## Importing data and libraries
```{r Imports}
#clear the working environment 
rm(list=ls())

#import libraries
library(ggplot2)

#import data
all_data <- readRDS("data/masked_outcome_measures.rds")
data_prb <- all_data$prb
data_mcsd <- all_data$mcsd
```

## 1. Understanding the structure of the data 
+ #### Data summaries\

```{r}
#general summary of the data
summary(data_prb)
summary(data_mcsd)
```
Columns **a** through **g** are the different conditions of the simulation. An interesting thing to note is that condition **1** of parameter **a** has only 540 occurrences and condition **8** of parameter **b** has 1620.

+ #### Number of unique conditions for each parameter:\
```{r}
apply(data_prb[1:7], 2, function(x) length(unique(x)))
```

The analysis will begin with the PRB because it is necessary to understand which conditions are biased in order to make cross-references with the MCSD.

+ #### Boxplots with condition on the x-axis and PRB on y-axis for int, x, z1:\
```{r}
#boxplots for prb with the different conditions for x
for (i in 1:length(data_prb[1:7])){
  print(ggplot(data_prb, aes(x=data_prb[, i], y=x)) + geom_boxplot() + xlab(colnames(data_prb[i])))
}
```
```{r}
#boxplots for prb with the different conditions for int
for (i in 1:length(data_prb[1:7])){
  print(ggplot(data_prb, aes(x=data_prb[, i], y=int)) + geom_boxplot() + xlab(colnames(data_prb[i])))
}
```
```{r}
#boxplots for prb with the different conditions for z1
for (i in 1:length(data_prb[1:7])){
  print(ggplot(data_prb, aes(x=data_prb[, i], y=z1)) + geom_boxplot() + xlab(colnames(data_prb[i])))
}
```
Parameter **a**: The largest PRB for the slope and intercept is in condition **3** and in condition **1** for the auxiliary variable.\

Parameter **b**: The mean PRB for *x* in all conditions is almost identical and below zero. The situation is similar for *int*, except for condition **8** - The 75th percentile range is higher and there are more outliers. The mean and 75th percentile range for **z1** in condition **8** is higher and the mean is lower than the other conditions. 

Parameter **c**: The highest mean PRB for **x** is in condition **3**. For **int** the mean is the same across conditions, however the IQR increases. Same goes for **z1**, although the increase is not as drastic. 

Parameter **d**: The IQR decreases for **x** until condition **4**. For **int** the IQR decreases drastically. For **z1** the situation is more interesting - The IQR increases until condition **3** and then decreases again. 

Parameter **e**: The IQR decreases in each condition for **x**. For **int** and **z1** the PRB distribution looks identical across conditions, except for the many outliers in condition **1** for **z1**. 

Parameter **f**: There seems to be no visible difference across the conditions for all three parameters. 

Parameter **g**: The IQR becomes smaller for **int** and **z1**.  Almost no difference in conditions for **x**. 

```{r}


```





